---
title: "Single Cell RNAseq Part 6"
author: "Bioinformatics Core"
output:
    html_document:
      keep_md: TRUE
---

Reminder of samples

* UCD_Adj_VitE
* UCD_Supp_VitE
* UCD_VitE_Def

## Load libraries
```{r, warning=FALSE,error=FALSE,message=FALSE}
library(Seurat)
library(ggplot2)
```

## Load the Seurat object
```{r}
load("clusters_seurat_object.RData")
experiment.merged
Idents(experiment.merged) <- "RNA_snn_res.0.5"
```


#0. Setup
Load the final Seurat object, load libraries (also see additional required packages for each example)
```{r, echo = F}
options(width = 450)
```

#1. DE With Single Cell Data Using Limma
For differential expression using models more complex than those allowed by FindAllMarkers(), data from Seurat may be used in limma (https://www.bioconductor.org/packages/devel/bioc/vignettes/limma/inst/doc/usersguide.pdf)

We illustrate by comparing sample 1 to sample 2 within cluster 0:
```{r}
library(limma)
cluster0 <- subset(experiment.merged, idents = '0')
expr <- as.matrix(GetAssayData(cluster0))

# Filter out genes that are 0 for every cell in this cluster
bad <- which(rowSums(expr) == 0)
expr <- expr[-bad,]

mm <- model.matrix(~0 + orig.ident, data = cluster0@meta.data)
fit <- lmFit(expr, mm)  
head(coef(fit)) # means in each sample for each gene
contr <- makeContrasts(orig.identUCD_Supp_VitE - orig.identUCD_Adj_VitE, levels = colnames(coef(fit)))
tmp <- contrasts.fit(fit, contrasts = contr)
tmp <- eBayes(tmp)
topTable(tmp, sort.by = "P", n = 20) # top 20 DE genes
```
* logFC: log2 fold change (UCD_Supp_VitE/UCD_Adj_VitE)
* AveExpr: Average expression, in log2 counts per million, across all cells included in analysis (i.e. those in cluster 0)
* t: t-statistic, i.e. logFC divided by its standard error
* P.Value: Raw p-value from test that logFC differs from 0
* adj.P.Val: Benjamini-Hochberg false discovery rate adjusted p-value

The limma vignette linked above gives more detail on model specification.

# 2. Gene Ontology (GO) Enrichment of Genes Expressed in a Cluster
```{r}
library(topGO)
# install org.Mm.eg.db from Bioconductor if not already installed (for mouse only)
cluster0 <- subset(experiment.merged, idents = '0')
expr <- as.matrix(GetAssayData(cluster0))
# Select genes that are expressed > 0 in at least 75% of cells (somewhat arbitrary definition)
n.gt.0 <- apply(expr, 1, function(x)length(which(x > 0)))
expressed.genes <- rownames(expr)[which(n.gt.0/ncol(expr) >= 0.75)]
all.genes <- rownames(expr)

# define geneList as 1 if gene is in expressed.genes, 0 otherwise
geneList <- ifelse(all.genes %in% expressed.genes, 1, 0)
names(geneList) <- all.genes

# Create topGOdata object
	GOdata <- new("topGOdata",
		ontology = "BP", # use biological process ontology
		allGenes = geneList,
		geneSelectionFun = function(x)(x == 1),
              annot = annFUN.org, mapping = "org.Mm.eg.db", ID = "symbol")
# Test for enrichment using Fisher's Exact Test
	resultFisher <- runTest(GOdata, algorithm = "elim", statistic = "fisher")
	GenTable(GOdata, Fisher = resultFisher, topNodes = 20, numChar = 60)

```
* Annotated: number of genes (out of all.genes) that are annotated with that GO term
* Significant: number of genes that are annotated with that GO term and meet our criteria for "expressed"
* Expected: Under random chance, number of genes that would be expected to be annotated with that GO term and meeting our criteria for "expressed"
* Fisher: (Raw) p-value from Fisher's Exact Test

#3. Weighted Gene Co-Expression Network Analysis (WGCNA)
WGCNA identifies groups of genes ("modules") with correlated expression.
WARNING: TAKES A LONG TIME TO RUN
```{r, warning = F}
library(WGCNA)
options(stringsAsFactors = F)
datExpr <- t(as.matrix(GetAssayData(experiment.merged)))[,VariableFeatures(experiment.merged)]  # only use variable genes in analysis

net <- blockwiseModules(datExpr, power = 10,
  corType = "bicor", # use robust correlation
	networkType = "signed", minModuleSize = 10,
	reassignThreshold = 0, mergeCutHeight = 0.15,
	numericLabels = F, pamRespectsDendro = FALSE,
	saveTOMs = TRUE,
	saveTOMFileBase = "TOM",
	verbose = 3)

table(net$colors)
# Convert labels to colors for plotting
mergedColors = net$colors
# Plot the dendrogram and the module colors underneath
plotDendroAndColors(net$dendrograms[[1]], mergedColors[net$blockGenes[[1]]],
"Module colors",
dendroLabels = FALSE, hang = 0.03,
addGuide = TRUE, guideHang = 0.05)
```
Genes in grey module are unclustered.

What genes are in the "blue" module?
```{r}
colnames(datExpr)[net$colors == "blue"]
```

Each cluster is represented by a summary "eigengene".
Plot eigengenes for each non-grey module by clusters from Seurat:
```{r}
f <- function(module){
  eigengene <- unlist(net$MEs[paste0("ME", module)])
  means <- tapply(eigengene, Idents(experiment.merged), mean, na.rm = T)
  return(means)
}
modules <- c("blue", "brown", "green", "turquoise", "yellow")
plotdat <- sapply(modules, f)
matplot(plotdat, col = modules, type = "l", lwd = 2, xaxt = "n", xlab = "Seurat Cluster",
        ylab = "WGCNA Module Eigengene")
axis(1, at = 1:19, labels = 0:18, cex.axis = 0.8)
matpoints(plotdat, col = modules, pch = 21)
```

Can also plot the module onto the tsne plot
```{r}
blue.eigengene <- unlist(net$MEs[paste0("ME", "blue")])
names(blue.eigengene) <- rownames(datExpr)
experiment.merged$blue.eigengene <- blue.eigengene
FeaturePlot(experiment.merged, features = "blue.eigengene", cols = c("grey", "blue"))
```


## Get the next Rmd file
```{r, eval=FALSE}
download.file("https://raw.githubusercontent.com/ucdavis-bioinformatics-training/2019-single-cell-RNA-sequencing-Workshop-UCD_UCSF/master/scrnaseq_analysis/scRNA_Workshop-PART7.Rmd", "scRNA_Workshop-PART7.Rmd")
```

## Session Information
```{r}
sessionInfo()
```
